### 2021-02-17 docs/Draft/draft.tex
update docs

\newcommand{\true}{\texttt{true}}
\newcommand{\false}{\texttt{false}}
\newcommand{\cref}{\texttt{ref}}
\newcommand{\cfork}[1]{\texttt{fork}\{#1\}}
\newcommand{\cloop}[1]{\texttt{loop}_{#1}\,}
\newcommand{\cbreak}{\texttt{break}\,}
\newcommand{\ccontinue}{\texttt{continue}}
\newcommand{\cif}{\texttt{if}\,}
\newcommand{\cthen}{\,\texttt{then}\,}
\newcommand{\celse}{\,\texttt{else}\,}
\newcommand{\cseq}{\,;;\,}
\newcommand{\creturn}{\texttt{return}\,}
\newcommand{\ccall}{\texttt{call}\,}

@@

Below shows the syntax of the programming language we consider, which is an extension of the one in the original Iris paper.
We also take Rust language as a reference in the design of the language.
$$
\begin{aligned}
    v \in \textit{Val} ::=\, & () \sep z \sep \true \sep \false \sep l \sep \lambda x.e \sep \cdots \\
    e \in \textit{Expr} ::=\, & v \sep x \sep e_1(e_2) \sep \cref(e) \sep !e \sep e_1 \leftarrow e_2 \sep \cfork{e} \sep e_1 \cseq e_2 \sep \cloop{e} e \\
                        \sep\, & \cif e_1 \cthen e_2 \celse e_3 \sep \cbreak e \sep \ccontinue \sep \ccall e \sep \creturn e \sep \cdots \\
    K \in \textit{Ctx} ::=\, & \bullet \sep K(e) \sep v(K) \sep \cref(K) \sep !K \sep K \leftarrow e \sep v \leftarrow K \sep K \cseq e \sep \cloop{e} K \\
                        \sep\, & \cif K \cthen e_2 \celse e_3 \sep \cbreak K \sep \ccall K \sep \creturn K \sep \cdots
\end{aligned}
$$

To allow context binding, we use $\cloop{e} e$ to represent a loop statement with $e$ as its loop body.
$e$ serves as a parameter in the corresponding evaluation context for the loop command.
After the completion of evaluation for expression in $K$, we can still recover the complete loop body from the subscription of $\cloop{e}\!$.
We can wrap the conventional loop definition as $\texttt{loop}\, e \triangleq \cloop{e} e$.

The $\cbreak\!$ statement and $\ccontinue$ statement are scoped by $\cloop{e}\!$ and can only manipulate the control flow within the loop.
The sequence statement and if statement definitions are standard.

We also consider control flow related to function invocations.
To the resolve the problem of whether the $\creturn\!$ statement can penetrate unnamed functions defined with the $\lambda$ operator (e.g. which one of 1 and 2 should $(\lambda x. \creturn x)(1) \cseq 2$ evaluate to), we use $\ccall\!$ to scope the influence of $\creturn\!$.
We can still define recursive functions in lambda calculus with the $\creturn\!$ functioning properly.

Following is an example defining the factorial function, where we early return 1 when $n$ is 0 and follows it by an invocation to $f$ recursively. The expression $\ccall (\textit{FACT}(n))$ would return the value of $n!$.
Removing $\ccall\!$ in this definition will cause the evaluation of $\textit{FACT}(n)$ always goes into $\creturn 1$.
$$
\begin{aligned}
    \textit{FACT}' &\triangleq \lambda f. \lambda n. (\cif n = 0 \cthen \creturn 1) \cseq n \times (\ccall (f\, f\, (n - 1))) \\
    \textit{FACT} &\triangleq \textit{FACT}'\,\textit{FACT}'
\end{aligned}
$$
% \todo{Probabily an example}

% $\cbreak\!$ and $\ccontinue$ can not penetrate $\cloop{e}\!$, and $\creturn\!$ can not penetrate $\ccall\!$.
% A problem, $(\lambda x. \creturn x)(1) \cseq 3$ should evaluate to 1 or 3? In other words, whether the lambda expression is a function or a normal expression?

@@

\begin{figure}[h]
$$
\begin{aligned}
    (\cloop{e} (), \sigma) &\hred (\cloop{e} e, \sigma, \epsilon) && \\
    (\cloop{e} (\ccontinue), \sigma) &\hred (\cloop{e} e, \sigma, \epsilon) &&  \\
    (\cloop{e} (\cbreak v), \sigma) &\hred (v, \sigma, \epsilon) && \\
    (K[\cbreak v], \sigma) &\hred (\cbreak v, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\cbreak\!} \\
    (K[\ccontinue], \sigma) &\hred (\ccontinue, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\ccontinue} \\
\end{aligned}
$$
\caption{Head Reductions for Loop}
\label{fig:hred-loop}
\end{figure}

Figure~\ref{fig:hred-loop} shows head reductions for $\cloop{e}\!$, $\cbreak\!$, and $\ccontinue$.
\begin{itemize}
\item The first two rules says that when the evaluation of the loop body finishes (must evaluate to the unit) or it encounters a $\ccontinue$, it reset the expression in the hole to $e$, the complete loop body.
\item The third rule says when the loop encounters a $\cbreak\!$ in the hole, the evaluation for the entire loop finishes with the value carried by $\cbreak\!$ as its result.
\item The last two rules says $\cbreak\!$ and $\ccontinue$ can terminate the reduction of their evaluation context $K$, if $K$ is their penetrable context $\text{PenCtx}$ defined as:
$$
\begin{aligned}
    \pure{\cbreak\!} &\triangleq \textit{Cxt}\,\backslash\cloop{e}\!\backslash\ccall \\
    \pure{\ccontinue} &\triangleq \textit{Cxt}\,\backslash\cloop{e}\!\backslash\ccall
\end{aligned}
$$
where $\textit{Cxt}\,\backslash\cloop{e}\!\backslash\ccall\!$ represents the set of contexts without using the $\cloop{e}\!$ and $\ccall\!$ constructs.
This limits effects of $\cbreak\!$ and $\ccontinue$ to stay within the loop body and function body.
We forbid $K$ to be $\bullet$ to avoid infinite reductions.
\end{itemize}

\begin{figure}[h]
$$
\begin{aligned}
    (\ccall v, \sigma) &\hred (v, \sigma, \epsilon) \\
    (\ccall (\creturn v), \sigma) &\hred (v, \sigma, \epsilon) \\
    (K[\creturn v], \sigma) &\hred (\creturn v, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\creturn\!}
\end{aligned}
$$
\caption{Head Reductions for Function Invocation}
\label{fig:hred-func}
\end{figure}

Figure~\ref{fig:hred-func} shows head reductions for $\ccall\!$ and $\creturn\!$.
\begin{itemize}
\item The first two rules says that when the function body evaluates to a value or a $\creturn\!$ statement, the entire function invocation statement evaluates to the corresponding value.
\item Similar to $\cbreak\!$ and $\ccontinue$, $\creturn\!$ can terminate the evaluation of its penetrable context:
$$
\pure{\creturn\!} \triangleq \textit{Cxt}\,\backslash\ccall
$$
which limits $\creturn\!$ to affect control flow within the innermost $\ccall\!$.
\item The $\creturn v$ itself can not reduce to $v$ outside a function invocation to allow the adequacy theorem hold for our weakest precondition definition, which seems reasonable if we consider any program consists of only an invocation to the main function.
\end{itemize}

\paragraph{Remark.} Notice that there are 4 types of terminals for an expression in the reduction: $v$, $\cbreak v$, $\ccontinue$, and $\creturn v$.
They cannot be further reduced and correspond to 4 different postconditions in the weakest precondition we defined later.
$$
\text{terminals} \triangleq \textit{Val} \cup \{\cbreak v, \ccontinue, \creturn v \sep v \in \textit{Val}\}
$$

\begin{figure}[h]
$$
\begin{aligned}
    (\cif \true \cthen e_2 \celse e_3, \sigma) &\hred (e_2, \sigma, \epsilon) && \\
    (\cif \false \cthen e_2 \celse e_3, \sigma) &\hred (e_3, \sigma, \epsilon) && \\
    (v \cseq e, \sigma) &\hred (e, \sigma, \epsilon) && \\
    (\cfork{e}, \sigma) &\hred ((), \sigma, e) && \text{if $e$ is well-formed} \\
    &\cdots
\end{aligned}
$$
\caption{Head Reductions for Other Expressions}
\label{fig:hred-other}
\end{figure}

Figure~\ref{fig:hred-other} shows the head reduction of the if statement, the sequence statement, and the fork statement.
First three rules are standard, while the last one for fork is modified from the one in the original Iris paper, which requires the expression in the forked thread to be well-formed.
\begin{definition}
    An expression is well-formed if and only if every $\cbreak\!$ and $\ccontinue$ in it is scoped by a $\cloop{e}\!$ and every $\creturn\!$ in it is scoped by a $\ccall\!$.    
\end{definition}

% $$
% \begin{aligned}
%     (\cloop{e} (), \sigma) &\hred (\cloop{e} e, \sigma, \epsilon) && \\
%     (\cloop{e} (\cbreak v), \sigma) &\hred (v, \sigma, \epsilon) && \\
%     (\cloop{e} (\ccontinue), \sigma) &\hred (\cloop{e} e, \sigma, \epsilon) &&  \\
%     (K[\cbreak v], \sigma) &\hred (\cbreak v, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\cbreak\!} \\
%     (K[\ccontinue], \sigma) &\hred (\ccontinue, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\ccontinue} \\
%     (\ccall v, \sigma) &\hred (v, \sigma, \epsilon) \\
%     (\ccall (\creturn v), \sigma) &\hred (v, \sigma, \epsilon) \\
%     (K[\creturn v], \sigma) &\hred (\creturn v, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\creturn\!} \\
%     (\cif \true \cthen e_2 \celse e_3, \sigma) &\hred (e_2, \sigma, \epsilon) && \\
%     (\cif \false \cthen e_2 \celse e_3, \sigma) &\hred (e_3, \sigma, \epsilon) && \\
%     (v \cseq e, \sigma) &\hred (e, \sigma, \epsilon) && \\
%     (\cfork{e}, \sigma) &\hred ((), \sigma, e) && \text{if } \wellf{e} \\
%     &\cdots
% \end{aligned}
% $$

% where
% $$
% \begin{aligned}
%     \pure{\cbreak\!} &\triangleq \textit{Cxt}\,\backslash\cloop{e}\!\backslash\ccall \\
%     % \{K \sep \forall K_1, K_2. \text{ if } K = K_1[K_2] \text{ then } K_1 \neq \cloop{e} K_2 \text{ and } K_2 \in \pure{\cbreak\!}\} \\
%     \pure{\ccontinue} &\triangleq \textit{Cxt}\,\backslash\cloop{e}\!\backslash\ccall \\
%     \pure{\creturn\!} &\triangleq \textit{Cxt}\,\backslash\ccall \\
%     % \{K \sep \forall K_1, K_2. \text{ if } K = K_1[K_2] \text{ then } K_1 \neq \cloop{e} K_2 \text{ and } K_2 \in \pure{\ccontinue}\} \\
%     \wellf{e} &\text{ iff. } % TODO: definition of well-definedness
%     \end{aligned}
% $$

% There is a small problem that the expression need to terminate after \cbreak{} and \ccontinue{}, otherwise the evaluation gets stuck.

@@

We define the weakest precondition with multiple postconditions for different exit type as follows:
$$
\begin{aligned}
    \wpre{e}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} \triangleq& \quad
           (e \in \textit{Val} \land \upd \Phi_N(e)) \\
    & \lor (e = \cbreak v \land \upd \Phi_B(v)) \\
    & \lor (e = \ccontinue \land \upd \Phi_C()) \\
    & \lor (e = \creturn v \land \upd \Phi_R(v)) \\
    & \lor \biggl(\forall \sigma. e \notin \text{terminals} \land S(\sigma) \wand \upd \bigl(\cred(e, \sigma) \\
    & \quad \land \later
        \forall e', \sigma', \vec{e}_f. \bigl((e, \sigma) \tred (e', \sigma', \vec{e}_f)\bigr) \wand \upd \\
    & \quad\quad\quad \bigl(S(\sigma') * \wpre{e'}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} * \varhexstar_{e'\in\vec{e}_f} \wpre{e'}{\top}{\bot}{\bot}{\bot} \bigr)\bigr)\biggr)
\end{aligned}
$$

There are 5 cases in total:
\begin{itemize}
\item \textbf{Normal exit:} If the expression has evaluated to a value, then $\Phi_N$ should hold.
\item \textbf{Break exit:} If the expression has evaluated to $\cbreak\!$, then $\Phi_B$ should hold.
\item \textbf{Continue exit:} If the expression has evaluated to $\ccontinue$, then $\Phi_C$ should hold.
\item \textbf{Return exit:} If the expression has evaluated to $\creturn\!$, then $\Phi_R$ should hold.
\item \textbf{Preservation:} If the expression is a non-terminal, then for any program state, the expression is reducible, and after one step of reduction, the weakest precondition holds for the new expression and any forked thread can terminate normally (where the well-formedness come into effect).
\end{itemize}

% $$
% \cred(e) \triangleq e \notin \textit{Val}\, \cup \{\cbreak v, \ccontinue, \creturn v\}
% $$

% The adequacy theorem will have some problem w.r.t the \creturn statement.

\begin{figure}[h]
    $$
    \begin{aligned}
        \Phi_B(v) &\vdash \wpre{\cbreak v}{\bot}{\Phi_B}{\bot}{\bot} & \textsc{(wp-break)} \\
        \Phi_C() &\vdash \wpre{\ccontinue}{\bot}{\bot}{\Phi_C}{\bot} & \textsc{(wp-continue)} \\
        \Phi_R(v) &\vdash \wpre{\creturn v}{\bot}{\bot}{\bot}{\Phi_R} & \textsc{(wp-return)} \\
        \Box(I \wand{}\, \wpre{e}{\_. I}{\Phi_B}{\_. I}{\Phi_R}) * I &\vdash \wpre{(\cloop{e}e)}{\Phi_B}{\bot}{\bot}{\Phi_R} & \textsc{(wp-loop)} \\
        % \wpre{e}{v. \wpre{K[v]}{\Phi_N}{\Phi_B}{\Phi_C}}{\bot}{\bot} &\vdash \wpre{K[e]}{\Phi_N}{\Phi_B}{\Phi_C} & \textsc{(wp-bind-normal)} \\
        % K\in\pure{\cbreak\!} \land \wpre{e}{\bot}{\Phi_B}{\bot} &\vdash \wpre{K[e]}{\bot}{\Phi_B}{\bot} & \textsc{(wp-bind-break)} \\
        \begin{aligned}
            &(v = \true \wand \wpre{e_2}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R}) \\
            \land &(v = \false \wand \wpre{e_3}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R})
        \end{aligned} &\vdash \wpre{\cif v \cthen e_2 \celse e_3}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} & \textsc{(wp-if)} \\
        \begin{aligned}
            \texttt{wp}\,e\,& \progspec{\_.\wpre{e_2}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R}} \\
                            & \progspec{\Phi_B}\,\progspec{\Phi_C}\,\progspec{\Phi_R}
        \end{aligned} &\vdash \wpre{e_1 \cseq e_2}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} & \textsc{(wp-seq)} \\
        \wpre{e}{\Phi_N}{\bot}{\bot}{\Phi_R} &\vdash \wpre{\ccall e}{\Phi_N \lor \Phi_R}{\bot}{\bot}{\bot} & \textsc{(wp-call)} \\
        \begin{aligned}
            \texttt{wp}\,e\,& \progspec{v. \wpre{K[v]}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R}} \\
                            & \progspec{v. K \in \pure{\cbreak\!} \land \Phi_B(v)} \\
                            & \progspec{v. K \in \pure{\ccontinue} \land \Phi_C(v)} \\
                            & \progspec{v. K \in \pure{\creturn\!} \land \Phi_R(v)}
        \end{aligned} &\vdash \wpre{K[e]}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} & \textsc{(wp-bind)} \\
        &\cdots
    \end{aligned}
    $$
    \caption{Some Rules for Weakest Precondition}
    \label{fig:wp-rules}
\end{figure}

>>>>

\newcommand{\true}{\texttt{true}}
\newcommand{\false}{\texttt{false}}
\newcommand{\cmdref}{\texttt{ref}}
\newcommand{\cmdfork}[1]{\texttt{fork}\{#1\}}
\newcommand{\cmdloop}[1]{\texttt{loop}_{#1}\,}
\newcommand{\cmdbreak}{\texttt{break}\,}
\newcommand{\cmdcontinue}{\texttt{continue}}
\newcommand{\cmdreturn}{\texttt{return}\,}
\newcommand{\cmdif}{\texttt{if}\,}
\newcommand{\cmdthen}{\,\texttt{then}\,}
\newcommand{\cmdelse}{\,\texttt{else}\,}
\newcommand{\cmdseq}{\,;;\,}
\newcommand{\cmdcall}{\texttt{call}\,}

@@

Below shows the syntax of the programming language we consider, which is an extension of the one in the original Iris paper.
We also take Rust language as a reference in the design of the language.
$$
\begin{aligned}
    v \in \textit{Val} ::=\, & () \sep z \sep \true \sep \false \sep l \sep \lambda x.e \sep \cdots \\
    e \in \textit{Expr} ::=\, & v \sep x \sep e_1(e_2) \sep \cmdref(e) \sep !e \sep e_1 \leftarrow e_2 \sep \cmdfork{e} \sep e_1 \cmdseq e_2 \sep \cmdloop{e} e \\
                        \sep\, & \cmdif e_1 \cmdthen e_2 \cmdelse e_3 \sep \cmdbreak e \sep \cmdcontinue \sep \cmdcall e \sep \cmdreturn e \sep \cdots \\
    K \in \textit{Ctx} ::=\, & \bullet \sep K(e) \sep v(K) \sep \cmdref(K) \sep !K \sep K \leftarrow e \sep v \leftarrow K \sep K \cmdseq e \sep \cmdloop{e} K \\
                        \sep\, & \cmdif K \cmdthen e_2 \cmdelse e_3 \sep \cmdbreak K \sep \cmdcall K \sep \cmdreturn K \sep \cdots
\end{aligned}
$$

To allow context binding, we use $\cmdloop{e} e$ to represent a loop statement with $e$ as its loop body.
$e$ serves as a parameter in the corresponding evaluation context for the loop command.
After the completion of evaluation for expression in $K$, we can still recover the complete loop body from the subscription of $\cmdloop{e}\!$.
We can wrap the conventional loop definition as $\texttt{loop}\, e \triangleq \cmdloop{e} e$.

The $\cmdbreak\!$ statement and $\cmdcontinue$ statement are scoped by $\cmdloop{e}\!$ and can only manipulate the control flow within the loop.
The sequence statement and if statement definitions are standard.

We also consider control flow related to function invocations.
To the resolve the problem of whether the $\cmdreturn\!$ statement can penetrate unnamed functions defined with the $\lambda$ operator (e.g. which one of 1 and 2 should $(\lambda x. \cmdreturn x)(1) \cmdseq 2$ evaluate to), we use $\cmdcall\!$ to scope the influence of $\cmdreturn\!$.
We can still define recursive functions in lambda calculus with the $\cmdreturn\!$ functioning properly.

Following is an example defining the factorial function, where we early return 1 when $n$ is 0 and follows it by an invocation to $f$ recursively. The expression $\cmdcall (\textit{FACT}(n))$ would return the value of $n!$.
Removing $\cmdcall\!$ in this definition will cause the evaluation of $\textit{FACT}(n)$ always goes into $\cmdreturn 1$.
$$
\begin{aligned}
    \textit{FACT}' &\triangleq \lambda f. \lambda n. (\cmdif n = 0 \cmdthen \cmdreturn 1) \cmdseq n \times (\cmdcall (f\, f\, (n - 1))) \\
    \textit{FACT} &\triangleq \textit{FACT}'\,\textit{FACT}'
\end{aligned}
$$
% \todo{Probabily an example}

% $\cmdbreak\!$ and $\cmdcontinue$ can not penetrate $\cmdloop{e}\!$, and $\cmdreturn\!$ can not penetrate $\cmdcall\!$.
% A problem, $(\lambda x. \cmdreturn x)(1) \cmdseq 3$ should evaluate to 1 or 3? In other words, whether the lambda expression is a function or a normal expression?

@@

\begin{figure}[h]
$$
\begin{aligned}
    (\cmdloop{e} v, \sigma) &\hred (\cmdloop{e} e, \sigma, \epsilon) && \\
    (\cmdloop{e} (\cmdcontinue), \sigma) &\hred (\cmdloop{e} e, \sigma, \epsilon) &&  \\
    (\cmdloop{e} (\cmdbreak v), \sigma) &\hred (v, \sigma, \epsilon) && \\
    (K[\cmdbreak v], \sigma) &\hred (\cmdbreak v, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\cmdbreak\!} \\
    (K[\cmdcontinue], \sigma) &\hred (\cmdcontinue, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\cmdcontinue} \\
\end{aligned}
$$
\caption{Head Reductions for Loop}
\label{fig:hred-loop}
\end{figure}

Figure~\ref{fig:hred-loop} shows head reductions for $\cmdloop{e}\!$, $\cmdbreak\!$, and $\cmdcontinue$.
\begin{itemize}
\item The first two rules says that when the evaluation of the loop body finishes or it encounters a $\cmdcontinue$, it reset the expression in the hole to $e$, the complete loop body. We no longer require the body evaluate to unit, and such requirement can be met by definition $\texttt{loop}\, e \triangleq \cmdloop{e} (e ;; ())$.
\item The third rule says when the loop encounters a $\cmdbreak\!$ in the hole, the evaluation for the entire loop finishes with the value carried by $\cmdbreak\!$ as its result.
\item The last two rules says $\cmdbreak\!$ and $\cmdcontinue$ can terminate the reduction of their evaluation context $K$, if $K$ is their penetrable context $\text{PenCtx}$ defined as:
$$
\begin{aligned}
    \pure{\cmdbreak\!} &\triangleq \textit{Cxt}\,\backslash\cmdloop{e}\!\backslash\cmdcall \\
    \pure{\cmdcontinue} &\triangleq \textit{Cxt}\,\backslash\cmdloop{e}\!\backslash\cmdcall
\end{aligned}
$$
where $\textit{Cxt}\,\backslash\cmdloop{e}\!\backslash\cmdcall\!$ represents the set of contexts without using the $\cmdloop{e}\!$ and $\cmdcall\!$ constructs.
This limits effects of $\cmdbreak\!$ and $\cmdcontinue$ to stay within the loop body and function body.
We forbid $K$ to be $\bullet$ to avoid infinite reductions.
\end{itemize}

\begin{figure}[h]
$$
\begin{aligned}
    (\cmdcall v, \sigma) &\hred (v, \sigma, \epsilon) \\
    (\cmdcall (\cmdreturn v), \sigma) &\hred (v, \sigma, \epsilon) \\
    (K[\cmdreturn v], \sigma) &\hred (\cmdreturn v, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\cmdreturn\!}
\end{aligned}
$$
\caption{Head Reductions for Function Invocation}
\label{fig:hred-func}
\end{figure}

Figure~\ref{fig:hred-func} shows head reductions for $\cmdcall\!$ and $\cmdreturn\!$.
\begin{itemize}
\item The first two rules says that when the function body evaluates to a value or a $\cmdreturn\!$ statement, the entire function invocation statement evaluates to the corresponding value.
\item Similar to $\cmdbreak\!$ and $\cmdcontinue$, $\cmdreturn\!$ can terminate the evaluation of its penetrable context:
$$
\pure{\cmdreturn\!} \triangleq \textit{Cxt}\,\backslash\cmdcall
$$
which limits $\cmdreturn\!$ to affect control flow within the innermost $\cmdcall\!$.
\item The $\cmdreturn v$ itself can not reduce to $v$ outside a function invocation to allow the adequacy theorem hold for our weakest precondition definition, which seems reasonable if we consider any program consists of only an invocation to the main function.
\end{itemize}

\paragraph{Remark.} Notice that there are 4 types of terminals for an expression in the reduction: $v$, $\cmdbreak v$, $\cmdcontinue$, and $\cmdreturn v$.
They cannot be further reduced and correspond to 4 different postconditions in the weakest precondition we defined later.
$$
\text{terminals} \triangleq \textit{Val} \cup \{\cmdbreak v, \cmdcontinue, \cmdreturn v \sep v \in \textit{Val}\}
$$

\begin{figure}[h]
$$
\begin{aligned}
    (\cmdif \true \cmdthen e_2 \cmdelse e_3, \sigma) &\hred (e_2, \sigma, \epsilon) && \\
    (\cmdif \false \cmdthen e_2 \cmdelse e_3, \sigma) &\hred (e_3, \sigma, \epsilon) && \\
    (v \cmdseq e, \sigma) &\hred (e, \sigma, \epsilon) && \\
    (\cmdfork{e}, \sigma) &\hred ((), \sigma, e) && \text{if $e$ is well-formed} \\
    &\cdots
\end{aligned}
$$
\caption{Head Reductions for Other Expressions}
\label{fig:hred-other}
\end{figure}

Figure~\ref{fig:hred-other} shows the head reduction of the if statement, the sequence statement, and the fork statement.
First three rules are standard, while the last one for fork is modified from the one in the original Iris paper, which requires the expression in the forked thread to be well-formed.
\begin{definition}
    An expression is well-formed if and only if every $\cmdbreak\!$ and $\cmdcontinue$ in it is scoped by a $\cmdloop{e}\!$ and every $\cmdreturn\!$ in it is scoped by a $\cmdcall\!$.    
\end{definition}

% $$
% \begin{aligned}
%     (\cmdloop{e} (), \sigma) &\hred (\cmdloop{e} e, \sigma, \epsilon) && \\
%     (\cmdloop{e} (\cmdbreak v), \sigma) &\hred (v, \sigma, \epsilon) && \\
%     (\cmdloop{e} (\cmdcontinue), \sigma) &\hred (\cmdloop{e} e, \sigma, \epsilon) &&  \\
%     (K[\cmdbreak v], \sigma) &\hred (\cmdbreak v, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\cmdbreak\!} \\
%     (K[\cmdcontinue], \sigma) &\hred (\cmdcontinue, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\cmdcontinue} \\
%     (\cmdcall v, \sigma) &\hred (v, \sigma, \epsilon) \\
%     (\cmdcall (\cmdreturn v), \sigma) &\hred (v, \sigma, \epsilon) \\
%     (K[\cmdreturn v], \sigma) &\hred (\cmdreturn v, \sigma, \epsilon) && \text{if } K \neq \bullet \text{ and } K \in \pure{\cmdreturn\!} \\
%     (\cmdif \true \cmdthen e_2 \cmdelse e_3, \sigma) &\hred (e_2, \sigma, \epsilon) && \\
%     (\cmdif \false \cmdthen e_2 \cmdelse e_3, \sigma) &\hred (e_3, \sigma, \epsilon) && \\
%     (v \cmdseq e, \sigma) &\hred (e, \sigma, \epsilon) && \\
%     (\cmdfork{e}, \sigma) &\hred ((), \sigma, e) && \text{if } \wellf{e} \\
%     &\cdots
% \end{aligned}
% $$

% where
% $$
% \begin{aligned}
%     \pure{\cmdbreak\!} &\triangleq \textit{Cxt}\,\backslash\cmdloop{e}\!\backslash\cmdcall \\
%     % \{K \sep \forall K_1, K_2. \text{ if } K = K_1[K_2] \text{ then } K_1 \neq \cmdloop{e} K_2 \text{ and } K_2 \in \pure{\cmdbreak\!}\} \\
%     \pure{\cmdcontinue} &\triangleq \textit{Cxt}\,\backslash\cmdloop{e}\!\backslash\cmdcall \\
%     \pure{\cmdreturn\!} &\triangleq \textit{Cxt}\,\backslash\cmdcall \\
%     % \{K \sep \forall K_1, K_2. \text{ if } K = K_1[K_2] \text{ then } K_1 \neq \cmdloop{e} K_2 \text{ and } K_2 \in \pure{\cmdcontinue}\} \\
%     \wellf{e} &\text{ iff. } % TODO: definition of well-definedness
%     \end{aligned}
% $$

% There is a small problem that the expression need to terminate after \cmdbreak{} and \cmdcontinue{}, otherwise the evaluation gets stuck.

@@

We define the weakest precondition with multiple postconditions for different exit type as follows:
$$
\begin{aligned}
    \wpre{e}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} \triangleq& \quad
           (e \in \textit{Val} \land \upd \Phi_N(e)) \\
    & \lor (\exists v \in \textit{Val}. e = \cmdbreak v \land \upd \Phi_B(v)) \\
    & \lor (e = \cmdcontinue \land \upd \Phi_C()) \\
    & \lor (\exists v \in \textit{Val}. e = \cmdreturn v \land \upd \Phi_R(v)) \\
    & \lor \biggl(\forall \sigma. e \notin \text{terminals} \land S(\sigma) \wand \upd \bigl(\cred(e, \sigma) \\
    & \quad \land \later
        \forall e', \sigma', \vec{e}_f. \bigl((e, \sigma) \tred (e', \sigma', \vec{e}_f)\bigr) \wand \upd \\
    & \quad\quad \bigl(S(\sigma') * \wpre{e'}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} * \varhexstar_{e'\in\vec{e}_f} \wpre{e'}{\top}{\top}{\top}{\top} \bigr)\bigr)\biggr)
\end{aligned}
$$

There are 5 cases in total:
\begin{itemize}
\item \textbf{Normal exit:} If the expression has evaluated to a value, then $\Phi_N$ should hold.
\item \textbf{Break exit:} If the expression has evaluated to $\cmdbreak\!$, then $\Phi_B$ should hold.
\item \textbf{Continue exit:} If the expression has evaluated to $\cmdcontinue$, then $\Phi_C$ should hold.
\item \textbf{Return exit:} If the expression has evaluated to $\cmdreturn\!$, then $\Phi_R$ should hold.
\item \textbf{Preservation:} If the expression is a non-terminal, then for any program state, the expression is reducible, and after one step of reduction, the weakest precondition holds for the new expression and any forked thread can terminate normally (where the well-formedness come into effect).
\end{itemize}

% $$
% \cred(e) \triangleq e \notin \textit{Val}\, \cup \{\cmdbreak v, \cmdcontinue, \cmdreturn v\}
% $$

% The adequacy theorem will have some problem w.r.t the \cmdreturn statement.

\begin{figure}[h]
    $$
    \begin{aligned}
        \Phi_B(v) &\vdash \wpre{\cmdbreak v}{\bot}{\Phi_B}{\bot}{\bot} & \textsc{(wp-break)} \\
        \Phi_C() &\vdash \wpre{\cmdcontinue}{\bot}{\bot}{\Phi_C}{\bot} & \textsc{(wp-continue)} \\
        \Phi_R(v) &\vdash \wpre{\cmdreturn v}{\bot}{\bot}{\bot}{\Phi_R} & \textsc{(wp-return)} \\
        \Box(I \wand{}\, \wpre{e}{\_. I}{\Phi_B}{\_. I}{\Phi_R}) * I &\vdash \wpre{(\cmdloop{e}e)}{\Phi_B}{\bot}{\bot}{\Phi_R} & \textsc{(wp-loop)} \\
        % \wpre{e}{v. \wpre{K[v]}{\Phi_N}{\Phi_B}{\Phi_C}}{\bot}{\bot} &\vdash \wpre{K[e]}{\Phi_N}{\Phi_B}{\Phi_C} & \textsc{(wp-bind-normal)} \\
        % K\in\pure{\cmdbreak\!} \land \wpre{e}{\bot}{\Phi_B}{\bot} &\vdash \wpre{K[e]}{\bot}{\Phi_B}{\bot} & \textsc{(wp-bind-break)} \\
        \begin{aligned}
            &(v = \true \wand \wpre{e_2}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R}) \\
            \land &(v = \false \wand \wpre{e_3}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R})
        \end{aligned} &\vdash \wpre{\cmdif v \cmdthen e_2 \cmdelse e_3}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} & \textsc{(wp-if)} \\
        \begin{aligned}
            \texttt{wp}\,e\,& \progspec{\_.\wpre{e_2}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R}} \\
                            & \progspec{\Phi_B}\,\progspec{\Phi_C}\,\progspec{\Phi_R}
        \end{aligned} &\vdash \wpre{e_1 \cmdseq e_2}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} & \textsc{(wp-seq)} \\
        \wpre{e}{\Phi}{\bot}{\bot}{\Phi} &\vdash \wpre{\cmdcall e}{\Phi}{\bot}{\bot}{\bot} & \textsc{(wp-call)} \\
        \begin{aligned}
            \texttt{wp}\,e\,& \progspec{v. \wpre{K[v]}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R}} \\
                            & \progspec{v. K \in \pure{\cmdbreak\!} \land \Phi_B(v)} \\
                            & \progspec{v. K \in \pure{\cmdcontinue} \land \Phi_C(v)} \\
                            & \progspec{v. K \in \pure{\cmdreturn\!} \land \Phi_R(v)}
        \end{aligned} &\vdash \wpre{K[e]}{\Phi_N}{\Phi_B}{\Phi_C}{\Phi_R} & \textsc{(wp-bind)} \\
        &\cdots
    \end{aligned}
    $$
    \caption{Some Rules for Weakest Precondition}
    \label{fig:wp-rules}
\end{figure}
